\normalfont{A bird's-eye view on the habitability of exoplanets via statistical learning techniques}

### 2020-09-25 Manuscript/Manuscript.tex
PUB

\clearpage
\subsection{Support vector machines}
The basic idea of the support vector machines is to consider the input set \textbf{X} ($n\times P$, where n stands for the observations and p for the features) as an ensemble of n-points immersed in to a p-dimensional space. Furthermore each point is labelled only in two way \footnote{The approaches described here can be extended to a multiple classes problem as shown in \cite{james2013introduction}, however since in this work only two classes are allowed (habitable or not) this such implementation will not be reviewed}. Thus, with this representation, the classification problem is recast in to a partition of a hyperspace. The simple form by which an hyperspace can be partitioned is an hyperplane:

\begin{equation}
\beta_{0}+\beta_{1}x_{i1}+\beta_{1}x_{i2}+...+\beta_{p}x_{ip} > 0\quad if \quad y_{i}=1
\end{equation}
\begin{equation}
\beta_{0}+\beta_{1}x_{i1}+\beta_{1}x_{i2}+...+\beta_{p}x_{ip} < 0\quad if \quad y_{i}=0
\end{equation}

The problem now is how to choose the hyperplane: this is a first discriminant between the SVM methods, moreover anything guarantee that the boundary is linear and thus a simple hypoerplane can be used; this is a second discriminant between the SVM methods. It is worth nothing that in principle one can introduce a further dimension, as shown in Fig \ref{Kernel_methods} in order to have a linear hyperplane: this is the idea of the kernel methods (further details are provided in \cite{james2013introduction,marsland2015machine}. 
\begin{figure}[h]
\begin{center}
\includegraphics[width=0.9\textwidth]{Pic/kernel_trick.png}
\caption{The trick of the kernel methods: introducing an extra dimension they transform a non-linear boundary in to a linear one.  Image taken from \cite{kernel}}
\label{Kernel_methods}
\end{center}
\end{figure}

>>>>

\subsection{Support vector machines}
The basic idea of the support vector machines is to consider the input set \textbf{X} ($n\times P$, where n stands for the observations and p for the features) as an ensemble of n-points immersed in to a p-dimensional space. Furthermore each point is labelled only in two way \footnote{The approaches described here can be extended to a multiple classes problem as shown in \cite{james2013introduction}, however since in this work only two classes are allowed (habitable or not) this such implementation will not be reviewed}. Thus, with this representation, the classification problem is recast in to a partition of a hyperspace. The simple form by which an hyperspace can be partitioned is an hyperplane \cite{james2013introduction}: 
\beta_{0}+\beta_{1}x_{i1}+\beta_{1}x_{i2}+...+\beta_{p}x_{ip} > 0\quad if \quad y_{i}=1
\beta_{0}+\beta_{1}x_{i1}+\beta_{1}x_{i2}+...+\beta_{p}x_{ip} < 0\quad if \quad y_{i}=0
The problem now is how to choose the hyperplane: this is a first discriminant between the SVM methods, moreover anything guarantee that the boundary is linear and thus a simple hyperplane can be used; this is a second discriminant between the SVM methods. It is worth nothing that in principle one can introduce a further dimension, as shown in Fig \ref{Kernel_methods} in order to have a linear hyperplane: this is the idea of the kernel methods\cite{james2013introduction,marsland2015machine}. 
\includegraphics[width=0.9\textwidth]{Pic/kernel_trick.png}
\caption{The trick of the kernel methods: introducing an extra dimension they transform a non-linear boundary in to a linear one. Image taken from \cite{kernel}}
\label{Kernel_methods}
\paragraph{Maximal margin Classifier}
The maximal margin classifier approach consist in tracing the separating hyperplane so that it has the  farthest minimum distance with respect to the training observation or more formally \cite{james2013introduction}: 
\max\limits_{{\beta_{0},\beta_{1}...\beta_{p},M}} M
subject\;to\;\sum_{j=1}^{p}\beta_{j}^{2}=1
y_{i}\left( \beta_{0} + \beta_{1} x_{i1} + \beta_{2} x_{i2}+...+\beta x_{ip}\right)\geq M\quad\forall\; i=1,...,n
\paragraph{Support vector Classifier}
The maximal margin approach assumes that the separating hyperplane exists, however this is far from be the most common type of classification problem. To overcome to this point the support vector classifier: in this approach the margin is considered soft allowing some point to be on the wrong side of the margin or on the wrong side of the hyperplane. The tolerance threshold is fixed by a quantity that is called budget $C$, more formally \cite{james2013introduction}:
\max\limits_{{\beta_{0},\beta_{1}...\beta_{p},M}} M
subject\;to\;\sum_{j=1}^{p}\beta_{j}^{2}=1
y_{i}\left( \beta_{0} + \beta_{1} x_{i1} + \beta_{2} x_{i2}+...+\beta x_{ip}\right)\geq M\left(1-\epsilon_{i}\right)
\epsilon_{i}\geq 0 \sum_{i=1}^{n}\epsilon_{i}\leq C
It is worth noting that if the budged is large a low variance will be found but with the toll of high bias; on the other side if the budget is low a low bias will be obtained but with the toll of high variance. Thus the selection of this parameter is crucial.  
\paragraph{Support vector machines}
This is the overall generalization of the method previously described: in particular the idea is to go beyond to the linear boundary. For this purpose the in the general expression \cite{james2013introduction}
f(x)=\beta_{0}+\sum_{i\in S}\alpha_{i}K(x,x_{i}) 
the linear kernel \cite{james2013introduction}
K(x_{i},x_{i'})=\sum^{p}_{j=1}x_{ij}x_{i'j}
is replaced with other kernels, as reported in Fig \ref{SVM_kernels} such as the polynomial of degree d \cite{james2013introduction}
K(x_{i},x_{i'})=\left(1+\sum_{j=1}^{p}x_{ij}x_{i'j}\right)^{d}
or perhaps with a radial one \cite{james2013introduction}
K(x_{i},x_{i'})=exp\left(-\gamma\sum^{p}_{j=1}\left(x_{ij}-x_{i'j}\right)\right)^{2}
where $\gamma$ is a positive constant. 
\includegraphics[width=0.9\textwidth]{Pic/Kernels_Type.png}
\caption{A set of possible type of kernels for the SVM. Image taken from \cite{ker_svm}}
\label{SVM_kernels}
